---
title: "Challenge 1"
author: "Pablo Barbera, Dan Cervone"
date: "January 20, 2016"
output: html_document
---

Write commands that help you answer the following questions using the dataset `congress-terms.csv`. Try not to look at the previous code we used!

```{r}
head(congress)
weird <- c(which(congress$age %in% c("No","Yes")), 
           grep("-", congress$age), 
           which(!(congress$chamber %in% c("house", "senate"))))
cong <- congress[-weird,]
```


1. Make a time series plot showing Democratic/Republican control of congress over time. By varying color, line thickness, or other graphical properties, make this plot also convey the average age of the congressmen representing each part over time.

```{r}
party <- with(cong, tapply(party, congress, function(xx) {
  partytable <- table(xx)
  ncong <- sum(partytable)
  ndem <- as.numeric(partytable["D"])
  nrep <- as.numeric(partytable["R"])
  nother <- ncong-ndem-nrep
  return(list(ncong=ncong, ndem=ndem, nrep=nrep, nother=nother))
}))

age <- with(cong, tapply(as.numeric(age), congress, mean, na.rm=T))
age_party <- with(cong, tapply(as.numeric(age), list(congress, party), mean, na.rm=T))

partydf <- do.call(rbind,lapply(party, data.frame))

partydf$age <- age
partydf$age_dem <- age_party[,"D"]
partydf$age_rep <- age_party[,"R"]

partydf <- partydf[order(as.numeric(rownames(partydf))),]

partydf$congress <- as.numeric(rownames(partydf))

library(ggplot2)

partydf$demratio <- partydf$ndem/partydf$ncong
partydf$ageratio <- partydf$age_dem/partydf$age_rep

myplot <- ggplot(data=partydf, 
                 mapping=aes(x=congress, y=demratio, color=age, size=age))
myplot + geom_line()

```

2. Write a function that, for a particular congress (e.g. congress 113), shows a map of the party breakdown by state.

```{r}
party_state <- with(cong[cong$congress==113,], tapply(party, state, function(xx) {
  partytable <- table(xx)
  ncong <- sum(partytable)
  ndem <- as.numeric(partytable["D"])
  nrep <- as.numeric(partytable["R"])
  nother <- ncong-ndem-nrep
  return(list(ncong=ncong, ndem=ndem, nrep=nrep, nother=nother))
}))

party_statedf <- do.call(rbind, lapply(party_state, data.frame))
party_statedf$state <- rownames(party_statedf)

statedf <- data.frame(cbind(state.abb, tolower(state.name)))

states <- map_data("state") # long/lat information for all states
party_statedf$region <- statedf$V2[match(party_statedf$state, statedf$state.abb)]

stateplot <- merge(states, party_statedf, by="region", all.x=TRUE)

head(states)
statemap <- ggplot(states, aes(x=long, y=lat, group=group)) + 
  geom_polygon(fill="grey65", color="white", size = 0.2) + # state colors/borders
  theme_bw() + # no background grid
  coord_map("mercator") # flat projection from sphere of US area
statemap
```